\textsf{Are large language models superhuman chemists?}

### 2024-10-16 src/tex/appendix.tex
reading ease

\begin{itemize}
    \item \emph{End-to-end automation}. For model development, the evaluations must be run many times (e.g., on regular intervals of a training run).
    Approaches that rely on humans scoring the answers of a system\autocite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
    \item \emph{Careful validation by experts}. Manual curation is needed to minimize the number of incorrect or unanswerable questions.\autocite{northcutt2021pervasive}
    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\autocite{Frye_2023, Awg}
    \item \emph{Usable with models that support special treatment of molecules}. Some models, such as Galactica\autocite{taylor2022galactica}, use special tokenization or encoding procedures for molecules or equations.
    The benchmark system must encode the semantic meaning of various parts of the question or answer to support this.
    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or raw logits.
    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
    Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real-world chemistry, as well as higher-level university education, multiple-choice questions are seldom utilized.
    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond answering \gls{mcq}.
    \item \emph{Cover a diverse set of topics}. Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
\end{itemize}

@@

\subsection{Model performance}
We also evaluated the model performance on the entire \chembench corpus. 
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models. 
Note that this ranking differs from the one on the \enquote{tiny} subset.

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic.pdf}
    \caption{\textbf{Fraction of completely correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic}
    \script{analyze_performance_per_source.py}
\end{figure}

@@

% \begin{figure}
%     \centering
%     \hspace*{-1cm}
%     \includegraphics{figures/reading_ease_vs_model_performance.pdf}
%     \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not. We do not observe a clear correlation between the reading ease of the questions and the performance of the models. }
%     \script{reading_ease_vs_model_performance.py}
%     \label{fig:reading_ease_vs_model_performance}
% \end{figure}

>>>>

\begin{itemize}
    \item \emph{End-to-end automation}. For model development, the evaluations must be run many times (e.g., on regular intervals of a training run).
    Approaches that rely on humans scoring the answers of a system\autocite{Schulze_Balhorn_2024, ai4science2023impact, castro2023large} can thus not be used.
    \item \emph{Careful validation by experts}. Manual curation is needed to minimize the number of incorrect or unanswerable questions.\autocite{northcutt2021pervasive}
    This is motivated by the observation that many widely used benchmarks are plagued by noisiness.\autocite{Frye_2023, Awg}
    \item \emph{Usable with models that support special treatment of molecules}. Some models, such as Galactica\autocite{taylor2022galactica}, use special tokenization or encoding procedures for molecules or equations.
    The benchmark system must encode the semantic meaning of various parts of the question or answer to support this.
    \item \emph{Usable with black box systems}. Many relevant systems do not provide access to model weights or raw logits.
    This might be the case because the systems are proprietary or because they involve not only \glspl{llm} but also external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
    Thus, a benchmark should not assume access to the raw model outputs but be able to operate on text completions.
    \item \emph{Probing capabilities beyond answering of \glspl{mcq}}. In real-world chemistry, as well as higher-level university education, multiple-choice questions are seldom utilized.
    Yet, most benchmarking frameworks focus on the \gls{mcq} setting because of the ease of evaluation. Realistic evaluations must measure capabilities beyond answering \gls{mcq}.
    \item \emph{Cover a diverse set of topics}. Chemistry, as the \enquote{central science}, bridges multiple disciplines.\autocite{Aspuru_Guzik_2018} To even just approximate \enquote{chemistry capabilities} the topics covered by a chemistry benchmark must be very diverse.
    \item \emph{Cover diverse skills}. To holistically judge performance it is important to cover questions relying on reasoning, calculation, knowledge, intuition. 
    \item \emph{Cover a range of difficulty levels}. To allow for a continuous measure of improvement for a range of different (evolving) systems, a benchmark should cover a wide range of difficulty levels.
    \item \emph{Impossible to completely solve with current models}. A benchmark should contain questions that are impossible to solve with current models. If current models can solve all questions, the benchmark provides no useful signal.
\end{itemize}

@@

\subsection{Model performance}
We also evaluated the model performance on the entire \chembench corpus. 
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models. 
Note that this ranking differs from the one on the \enquote{tiny} subset

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic.pdf}
    \caption{\textbf{Fraction of correctly answered questions per data source.} The heatmap shows, in color, the fraction of questions answered correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none of the questions were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for our leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic}
    \script{analyze_performance_per_source.py}
\end{figure}

@@

\begin{figure}
    \centering
    \hspace*{-1cm}
    \includegraphics{figures/reading_ease_vs_model_performance.pdf}
    \caption{\textbf{Model performance as a function of reading ease.} The violin plots show the distribution of reading ease scores for questions answered completely correctly and those not. We do not observe a clear correlation between the reading ease of the questions and the performance of the models. }
    \script{reading_ease_vs_model_performance.py}
    \label{fig:reading_ease_vs_model_performance}
\end{figure}
